### 2013-05-08 doc/icml2013/algo.tex
related work

\begin{algorithm}[t]
  \caption{Spectral Experts}
  \label{algo:spectral-experts}
  \begin{algorithmic}[1]
    \INPUT Three datasets $\mathcal{D}_r = \{ (\xn{1}, \yn{1}), \cdots, (\xn{n}, \yn{n}) \}$ for $r = 1, 2, 3$;
    %regularization strengths $\lambda_n^{(2)} = \frac{c_2}{\sqrt{n}}$, $\lambda_n^{(3)} = \frac{c_3}{\sqrt{n}}$;
    regularization strengths $\lambda_n^{(2)}$, $\lambda_n^{(3)}$;
    observation noise variance $\E[\epsilon^2]$.
    \OUTPUT Parameters $\hat\theta = (\hat \pi, [\hat \beta_1 \mid \cdots \mid \hat \beta_k])$.
    \STATE Estimate compound parameters $M_2, M_3$ using \textbf{low-rank regression}:
    \begin{align}
      &\hat M_1 = \arg\min_{M_1} \label{eqn:estimateM1} \\
      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_1} (\innerp{M_1}{x} - y)^2, \nonumber \\
      &\hat M_2 = \arg\min_{M_2} \quad \lambda_n^{(2)} \|M_2\|_* + \label{eqn:estimateM2} \\
      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_2} (\innerp{M_2}{x\tp{2}} + \E[\epsilon^2] - y^2)^2, \nonumber \\
      &\hat M_3 = \arg\min_{M_3} \quad \lambda_n^{(3)} \|M_3\|_* + \label{eqn:estimateM3} \\
      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_3} (\innerp{M_3}{x\tp{3}} + 3 \E[\epsilon^2]\innerp{\hat M_1}{x} - y^3)^2. \nonumber
    \end{align}
    \STATE Estimate the parameters $\theta = (\pi, B)$ using \textbf{tensor factorization}:
    \begin{enumerate}
      \item [(a)] Compute whitening matrix $\hat W \in \Re^{d \times k}$ (such that $\hat W^\top
      \hat M_2 \hat W = I$) using SVD.
      \item [(b)] Compute the eigenvalues $\{\hat a_h\}_{h=1}^k$
      and eigenvectors $\{\hat v_h\}_{h=1}^k$
      of the whitened tensor $\hat M_3(\hat W, \hat W, \hat W) \in \Re^{k \times k \times k}$
      by using the robust tensor power method.
    \item [(c)] Set parameter estimates $\hat\pi_h = \frac{1}{\hat a_h^2}$
    and $\hat\beta_h = (\hat W^{\top})^\dagger (\hat a_h \hat v_h)$.
    \end{enumerate}
  \end{algorithmic}
\end{algorithm}

@@

\paragraph{Remarks}

% Spectral
In recent years, there has a been substantial interest in ``spectral'' methods
for learning latent-variable models.  One line of work has
focused on observable operator models \cite{hsu09spectral}
in which the true parameters are not recovered, but a reparametrization is used
which enables prediction.
Another line of work is based on the method of moments and uses eigendecomposition of a certain tensor
to recover the parameters \cite{anandkumar12lda,anandkumar12moments}.
Our work extends this line of work into the space of discriminative models,
which requires regression to obtain the desired tensor.

@@

>>>>

\begin{algorithm}[t]
  \caption{Spectral Experts}
  \label{algo:spectral-experts}
  \begin{algorithmic}[1]
    \INPUT Datasets $\mathcal{D}_p = \{ (\xn{1}, \yn{1}), \cdots, (\xn{n}, \yn{n}) \}$ for $p = 1, 2, 3$;
    %regularization strengths $\lambda_n^{(2)} = \frac{c_2}{\sqrt{n}}$, $\lambda_n^{(3)} = \frac{c_3}{\sqrt{n}}$;
    regularization strengths $\lambda_n^{(2)}$, $\lambda_n^{(3)}$;
    observation noise moments $\E[\epsilon^2], \E[\epsilon^3]$.
    \OUTPUT Parameters $\hat\theta = (\hat \pi, [\hat \beta_1 \mid \cdots \mid \hat \beta_k])$.
    \STATE Estimate compound parameters $M_2, M_3$ using \textbf{low-rank regression}:
    \begin{align}
      &\hat M_1 = \arg\min_{M_1} \label{eqn:estimateM1} \\
      &\quad\sum_{(x,y) \in \sD_1} (\innerp{M_1}{x} - y)^2, \nonumber \\
      &\hat M_2 = \arg\min_{M_2} \quad \lambda_n^{(2)} \|M_2\|_* + \label{eqn:estimateM2} \\
      &\quad\sum_{(x,y) \in \sD_2} (\innerp{M_2}{x\tp{2}} + \E[\epsilon^2] - y^2)^2, \nonumber \\
      &\hat M_3 = \arg\min_{M_3} \quad \lambda_n^{(3)} \|M_3\|_* + \label{eqn:estimateM3} \\
      &\quad\sum_{(x,y) \in \sD_3} (\innerp{M_3}{x\tp{3}} + 3 \E[\epsilon^2]\innerp{\hat M_1}{x} + \E[\epsilon^3] - y^3)^2. \nonumber
    \end{align}
    \STATE Estimate parameters $\theta = (\pi, B)$ using \textbf{tensor factorization}:
    \begin{enumerate}
      \item [(a)] Compute whitening matrix $\hat W \in \Re^{d \times k}$ (such that $\hat W^\top
      \hat M_2 \hat W = I$) using SVD.
      \item [(b)] Compute eigenvalues $\{\hat a_h\}_{h=1}^k$
      and eigenvectors $\{\hat v_h\}_{h=1}^k$
      of the whitened tensor $\hat M_3(\hat W, \hat W, \hat W) \in \Re^{k \times k \times k}$
      by using the robust tensor power method.
    \item [(c)] Return parameter estimates $\hat\pi_h = \hat a_h^{-2}$
    and $\hat\beta_h = (\hat W^{\top})^\dagger (\hat a_h \hat v_h)$.
    \end{enumerate}
  \end{algorithmic}
\end{algorithm}

@@

\paragraph{Related work}

% Spectral
In recent years, there has a been a surge of interest in ``spectral'' methods
for learning latent-variable models.  One line of work has
focused on observable operator models \cite{hsu09spectral,song10kernel,parikh12spectral,cohen12pcfg,balle11transducer,balle12automata}
in which a reparametrization of the true parameters are recovered,
which suffices for prediction and density estimation.
Another line of work is based on the method of moments and uses eigendecomposition of a certain tensor
to recover the parameters \cite{anandkumar12moments,anandkumar12lda,hsu12identifiability,hsu13spherical}.
Our work extends this second line of work to models that
require regression to obtain the desired tensor.

@@

% 
%Previous work has focused on using spectral methods to learn finite state automata and transducers.
Other work has focused on learning discriminative models,
notably \citet{balle11transducer} for finite state transducers (functions from strings to strings),
and \citet{balle12automata} for weighted finite state automata (functions from strings to real numbers).
Similar to Spectral Experts,
\citet{balle12automata} used a two-step approach,
where convex optimization is first used to estimate moments (the Hankel matrix in their case),
after which these moments are subjected to spectral decomposition.  
However, these methods are developed in the observable operator framework, whereas we consider parameter estimation.
